---
title: "SMDS Homework - Block 4"
author: "C. Dorigo, R. Marvulli, A. Scardoni, L. Arrighi | Group 'C'"
date: "29th April 2020"
output:
  html_document:
    toc: yes
  beamer_presentation:
    highlight: tango
  include: null
  ioslides_presentation:
    highlight: tango
  pdf_document:
    highlight: tango
    keep_tex: yes
    toc: yes
  slide_level: 2
  slidy_presentation:
    fig.height: 3
    fig.width: 4
    highlight: tango
header-includes:
- \usepackage{color}
- \usepackage{graphicx}
- \usepackage{grffile}
- \definecolor{Purple}{HTML}{911146}
- \definecolor{Orange}{HTML}{CF4A30}
- \setbeamercolor{alerted text}{fg=Orange}
- \setbeamercolor{frametitle}{bg=Purple}
institute: University of Trieste, SISSA, ICTP, University of Udine
graphics: yes
fontsize: 10pt
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.align = 'center', warning=FALSE, message=FALSE, fig.asp=0.625, dev='png', global.par = TRUE, dev.args=list(pointsize=10), fig.path = 'figs/')
library(MASS)
```

```{r setup, include=FALSE}
library(knitr)
local({
  hook_plot = knit_hooks$get('plot')
  knit_hooks$set(plot = function(x, options) {
    paste0('\n\n----\n\n', hook_plot(x, options))
  })
})
```

# DAAG EXERCISES {.tabset}

## Exercise 1 

Exercise text here.

### Solution

```{r lab_0x, code = readLines("src/lab_0x.R"), echo=TRUE}
```

Solution here.

### Observations

Comments here.


## Chapter 8 - Exercise 1 

The following table shows numbers of occasions when inhibition (i.e., no flow of current across a membrane) occurred within $120$ s, for different concentrations of the protein peptide-C (data are used with the permission of Claudia Haarmann, who obtained these data in the course of her PhD research). The outcome $\texttt{yes}$ implies that inhibition has occurred.

Use logistic regression to model the probability of inhibition as a function of protein concentration.

### Solution

```{r}
conc <- c(.1, .5, 1, 10, 20, 30, 50, 70, 80, 100, 150)
```

here's the conc

```{r}
no <- c(7, 1, 10, 9, 2, 9, 13, 1, 1, 4, 3)
yes <- c(0, 0, 3, 4, 0, 6, 7, 0, 0, 1 ,7)
```

here's no+yes
let's sum

```{r}
n <- no + yes
```

plot

```{r}
plot(log(conc), log((yes+0.5)/(no+0.5)))
```

### Observations

Comments here.

